---
title: "Tutorial 7 - Model Selection."
author: "<INSERT YOUR NAME>"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to the seventh tutorial of the Regression and Simulation methods module. This is the next script in developing your skills in R, whilst learning about model selection. Throughout this notebook we will start to consider how to implement different procedures necessary for model selection.

### Exercise 0

Throughout remember we will need tidyverse, go ahead and do this as your first task.

Your Answer:

```{r Exercise0, message=FALSE}
library(tidyverse)
```

------------------------------------------------------------------------

### Exercise 1

In this tutorial we will follow the suit of the lecture R script and try and apply this to the data set that we explored the previous week. Below we quickly clean the data in the same manner as the previous week.

For our purposes we will be using a 'cleaned' set of this data. It is called, '31days.csv' Go ahead and load this in below.

```{r Exercise1}
ThirtyOneDays <- read_csv("/Users/chrisoldnall/Library/Mobile Documents/com~apple~CloudDocs/Teaching/SMSTC_RegAndSim_2023/Week 6/31days.csv")

HealthBoards <- c('S08000020','S08000022', 'S08000025', 'S08000026', 'S08000030', 'S08000028', 'S08000016','S08000017','S08000029','S08000024', 'S08000015','S08000019','S08000031','S08000032')
CancerAreas <- c('NCA', 'NCA', 'NCA', 'NCA', 'NCA', 'NCA', 'SCAN', 'SCAN', 'SCAN', 'SCAN', 'WOSCAN', 'WOSCAN', 'WOSCAN', 'WOSCAN')
df_mapping <- data.frame(HealthBoards = HealthBoards, CancerAreas = CancerAreas)

ThirtyOneDays <- ThirtyOneDays %>%
  left_join(df_mapping, by=c("HB" = "HealthBoards"))

ThirtyOneDays_Aggregated <- aggregate(cbind(NumberOfEligibleReferrals31DayStandard=ThirtyOneDays$NumberOfEligibleReferrals31DayStandard,  NumberOfEligibleReferralsTreatedWithin31Days=ThirtyOneDays$NumberOfEligibleReferralsTreatedWithin31Days, PopSize = ThirtyOneDays$PopSize), by=list(Quarter = ThirtyOneDays$Quarter, Date=ThirtyOneDays$Date, Pandemic=ThirtyOneDays$Pandemic, JustPandemic=ThirtyOneDays$JustPandemic, Pandemic_cat=ThirtyOneDays$Pandemic_cat), FUN=sum)
```

------------------------------------------------------------------------

### Exercise 2

DOD Backward Selection: In a similar way to the lecture code, perform backward selection with the variable of number of patients treated as the outcome variable. Ensure you comment on how this selection is done.

```{r Exercise2}

```

------------------------------------------------------------------------

### Exercise 3

Wald Test Backward Selection: Now perform the same test over again, but this time perform the Wald based selection procedure. Ensure you comment on how this selection is done, importantly noting if there are any differences in procedure outcomes. If so, what does this mean?

```{r Exercise3}

```

------------------------------------------------------------------------

### Exercise 4

AIC/BIC Selection Procedure: Finally conduct either AIC or BIC. If you do both think about how you would interpret conflicting results. Is one 'better' than the other.

```{r Exercise4}

```

------------------------------------------------------------------------

Now that you have completed this notebook go ahead and ensure you have completed all the tasks from the Week 7 SMSTC exercises.